### 2020-10-26 doc/LG5_errata/LG5_errata.Rmd
Adding LG5 model errata docs

---
title: "LG5 Function Errata"
author: "Albert Y. Kim"
date: "Last updated on `r Sys.Date()`"
output:
  pdf_document:
    number_sections: yes
  html_document:
    df_print: kable
    highlight: tango
    number_sections: yes
    theme: cosmo
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, echo = FALSE)
library(tidyverse)
```

# LG5 model formula typo

The formula is written as follows in [McMahon and Parker 2015 Eq 1 (page 245)](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4314258/). Note how in the numerator the $L$'s cancel out, leaving only $K$ being included in the formula:

$$
dbh = \frac{L + (K-L)}{1 + 1/\theta*\exp(-r(doy - doy_{ip})/\theta)^{\theta}}
$$
The formula is written as follows in the D'Orangeville et al. draft manuscript (page 9). Note the numerator is different than McMahon and Parker's formulation:

$$
dbh = \frac{K-L}{1 + 1/\theta*\exp(-r(doy - doy_{ip})/\theta)^{\theta}}
$$
However the `lg5.pred()` function from the [`RDendrom`
package](https://github.com/seanmcm/RDendrom/blob/master/R/DENDRO_BASE_FUNCTIONS.R#L544) computes DBH as follows:

```{r, echo = TRUE}
lg5.pred <- function(params, doy) {
  L <- params[1] # min(dbh, na.rm = T)
  K <- params[2]
  doy.ip <- params[3]
  r <- params[4]
  theta <- params[5]
  dbh <- vector(length = length(doy))
  dbh <- L + ((K - L) / (1 + 1 / theta * exp(-(r * (doy - doy.ip) / theta))^theta))
  return(dbh)
}
```

which corresponds to the following formula:

$$
dbh = L + \frac{K-L}{1 + 1/\theta*\exp(-r(doy - doy_{ip})/\theta)^{\theta}}
$$

Also note that the denominator of the LG5 function can be simplified since 
$$
\exp(-r(doy - doy_{ip})/\theta)^{\theta} = \exp(-r(doy - doy_{ip}))
$$

# $r$ parameter from LG5 model

I have a suspicion both these interpretations of the $r$ parameter of LG5 model are incorrect:

1. McMahon and Parker 2015 state just above Eq 1 (page 245): "the rate parameter $r$ describes the slope of the curve at the inflection point"
1. D'Orangeville et al. state (page 9): "$r$ represents the maximum growth rate" and "$doy_{ip}$ is the day of year when maximum growth rate occurs"

Rather, I believe that the correct formulation of the slope/the maximum growth rate is instead:

$$
r_{new} = \frac{K-L}{\left(1 + \frac{1}{\theta}\right)^2}\frac{r}{\theta}
$$

## Plot LG5 diameter and growth curves

Let's set arbitrarily chosen parameter values then

1. Use `lg5.pred()` to compute all diameters for all DOY = 1, ..., 365
1. Compute growth by taking differences of diameters

```{r, echo = TRUE}
L <- 13
K <- 14
doy.ip <- 175
r <- 0.05
theta <- 1
```

We then plot both the diameter curve as well as the growth curve, with `doy.ip` marked with a vertical dotted line. Observe how growth peaks at `doy.ip`.

```{r, fig.width=10}
params <- c(L, K, doy.ip, r, theta)
dbh_growth_values <- tibble(doy = seq(from = 1, to = 365)) %>%
  mutate(
    diameter = lg5.pred(params, doy),
    growth = diameter - lag(diameter)
  )

par(mfrow=c(1, 2))
plot(dbh_growth_values$doy, dbh_growth_values$diameter, type = "l", 
     main = "LG5 diameter", xlab = "doy", ylab = "diameter")
abline(v = doy.ip, lty = "dotted")
plot(dbh_growth_values$doy, dbh_growth_values$growth, type = "l", 
     main = "Growth curve", xlab = "doy", ylab = "growth")
abline(v = doy.ip, lty = "dotted")
```

## Original interpretations of $r$

1. McMahon and Parker 2015 state: "the rate parameter $r$ describes the slope of the curve at the inflection point". We mark the appropriate line with slope `r` in orange in the diameter plot. 
1. D'Orangeville et al. state: "$r$ represents the maximum growth rate" and "$doy_{ip}$ is the day of year when maximum growth rate occurs". Let's mark this maximum growth rate with a horizontal orange line in the growth plot.

```{r, fig.width=10}
# Compute diameter at doy.ip, then solve for intercept via y = mx + b 
y <- lg5.pred(params, doy.ip)
intercept <- y - r*doy.ip

par(mfrow=c(1, 2))
plot(dbh_growth_values$doy, dbh_growth_values$diameter, type = "l", 
     main = "LG5 diameter", xlab = "doy", ylab = "diameter")
abline(v = doy.ip, lty = "dotted")
abline(intercept, r, col = "orange")
plot(dbh_growth_values$doy, dbh_growth_values$growth, type = "l", 
     main = "Growth curve", xlab = "doy", ylab = "growth")
abline(v = doy.ip, lty = "dotted")
abline(h = r, col = "orange")
```

Observe

1. In the left plot how the orange line does not appear to be tangent to the LG5 curve at `doy.ip` and hence is not the slope at that point.
1. In the right plot $r$ = `r r` does not correspond to the growth rate at `doy.ip` which should be the max growth rate. Rather it appears to be around 0.012

## Proposed new $r$

Let's repeat the above with our proposed growth rate $r_{new}$, which is $\frac{d}{dx}LG5(x)$ evaluated at $x$ = `doy.ip`. See this [PDF](https://drive.google.com/file/d/1CMT2o1svTfbS7wQ_uJoWjqHbvDopkCZw/view?usp=sharing) for the analytic derivation of the derivative.

$$
r_{new} = \frac{K-L}{\left(1 + \frac{1}{\theta}\right)^2}\frac{r}{\theta}
$$

```{r, fig.width=10}
# Comupte new r, then find intercept
r_new <- ((K-L)/(1 + 1/theta)^2) * (r/theta)
intercept_new <- y - r_new*doy.ip

par(mfrow=c(1, 2))
plot(dbh_growth_values$doy, dbh_growth_values$diameter, type = "l", 
     main = "LG5 diameter", xlab = "doy", ylab = "diameter")
abline(v = doy.ip, lty = "dotted")
abline(intercept_new, r_new, col = "orange")
plot(dbh_growth_values$doy, dbh_growth_values$growth, type = "l", 
     main = "Growth curve", xlab = "doy", ylab = "growth")
abline(v = doy.ip, lty = "dotted")
abline(h = r_new, col = "orange")
```

Observe:

1. In the left plot how the orange line is now tangent to the LG5 curve at `doy.ip` and thus is the correct slope.
1. In the right plot $r_{new}$ = `r r_new` now corresponds to the max growth rate occurring at `doy.ip`.